---
title: "Homework 5"
output: html_notebook
---

(1) Do opposites attract? Three Duke undergraduates conducted a study of this question in 1993. They asked
questions of 28 female undergraduates concerning their own personalities. These included the following five
questions: Are you 1. introverted, 2. easy-going, 3. studious, 4. serious, 5. gregarious? They asked the same five questions about the respondent’s ideal mates. Take “opposites attract” to mean that the respondent answers
differently for their ideal mates than they do for themselves in at least two of the five questions. In this sample of 28, only 8 were attracted to opposites. Deeming this to be a random sample of Duke undergraduates,
report and interpret in context a 90% posterior credible interval for the population proportion of those
attracted to opposites. Use a beta(1, 1) prior density.

```{r}
alpha <- 1 ; beta <- 1
success <- 8 ; length <- 28
alphastar <- alpha + success
betastar <- length - success + beta
```

```{r}
creds <- qbeta(c(0.05, 0.95),9,21)
paste("We are 90% sure that the posterior probability of the proportion of Duke undergraduate females attracted to opposites is between",round(creds[1],3),"and",round(creds[2],3))
```


(2) A study by Charles J. Graham and others at Arkansas Children’s Hospital in Little Rock addressed the question of whether left-handed children are more accident prone than right-handed children. Of 267 children between the ages of 6 and 18 who were admitted to a pediatric emergency room for trauma, 44 of them (or 16.5%) were indeed left-handed. The investigators claimed that about 10% of all children are left-handed. (Indeed, this was about the proportion of children who were admitted to the same emergency room for nontrauma reasons.) Consider the population of all pediatric trauma patients. Let θ represent the proportion of all pediatric trauma patients who are left-handed. Assume the beta(2, 18) prior density for θ.

```{r}
alpha <- 2 ; beta <- 18
success <- 44 ; length <- 267
alphastar <- alpha + success
betastar <- length - success + beta
```


(a) For the Beta(2, 18) prior density, what is the a priori expected value of θ?

```{r}
alpha/(alpha + beta)
```


(b) Find and interpret a 95% posterior credibility interval for the population proportion of left-handers among
all pediatric trauma patients.

```{r}
creds <- qbeta(c(0.025, 0.975),alphastar,betastar)
paste("We are 95% sure that the posterior probability of the proportion of all pediatric trauma patients who are left-handed is between",round(creds[1], 4),"and",round(creds[2],4))
```

(c) Relying on the credible interval in part b, is it reasonable to claim that left-handed and right-handed
children are equally accident prone, or is there substantial evidence that left-handed children are more
accident prone? Explain your answer.


<b>It is not reasonable to assume that left-handed children are less accident prone than right-handed children as the credible interval for the proportion of left-handed children is above .10 which is the proportion of children who were admitted to the same emergency room for nontrauma reasons.</b>

(d) Find the posterior probability that θ is greater than 10%.

```{r}
1- pbeta(.1,alphastar,betastar)
```

(e) What is the posterior expected value of θ?

```{r}
alphastar/(alphastar+betastar)
```

(3) In a Newsweek poll taken on February 18-19, 1993, of 753 adults who were asked, “Would you favor additional
taxes to pay for reforming and expanding health care in the United States?” 65% (or 489) said “yes.” Assuming
a beta(20,20) prior density, report and interpret in context a 95% credible interval for the population proportion who would have answered “yes.”

```{r}
alpha <- 20 ; beta <- 20
success <- 489 ; length <- 753
alphastar <- alpha + success
betastar <- length - success + beta
```

```{r}
creds <- qbeta(c(0.025, 0.975),alphastar,betastar)
paste("We are 95% sure that the posterior probability of the proportion who would pay to reform and expand health care in the US is between",round(creds[1], 4),"and",round(creds[2],4))
```

(4) Repeat the previous question, but this time using a beta(1,1) prior density instead of a beta(20,20) prior
density. Also comment on how sensitive the resulting intervals are to the choice of which of these two priors
was used.

```{r}
alpha <- 1 ; beta <- 1
success <- 489 ; length <- 753
alphastar <- alpha + success
betastar <- length - success + beta
```

```{r}
creds <- qbeta(c(0.025, 0.975),alphastar,betastar)
paste("We are 95% sure that the poseterior probability of the proportion who would pay to reform and expand health care in the US is between",round(creds[1], 4),"and",round(creds[2],4))
```

<b> It doesn't seem that the resulting intervals are too sensitive based on the priors used as it only changed the intervals by about 1%. This is because there is so much more data than our confidence on the priors</b>

(5) In a documentary film A Private Universe, only 2 of 23 gowned graduates at a Harvard University commencement were able to correctly answer the question: “Why is it hotter in summer than in winter?” In view of the results of the survey, it is difficult for me to believe that these 23 are a random sample of all Harvard graduates, but assume that they are for the sake of calculation. Assume a beta(1, 1) density for θ ≡ the proportion of Harvard graduates who would answer this question correctly.

(a) What is the posterior distribution of θ?

```{r}
alpha <- 1 ; beta <- 1
success <- 2 ; length <- 23
alphastar <- alpha + success
betastar <- length - success + beta
paste("Beta(",alphastar,",",betastar,")", sep = "")
```

(b) Plot the prior (in gray) and the posterior (in black) densities on the same graph, and properly format
this graph (including a legend to identify the prior and posterior).

```{r}
xx <- seq(0,1,length.out = 1000)
prior <- dbeta(xx,1,1)
posterior <- dbeta(xx,alphastar,betastar)
plot(xx,prior,type = "l", col = "light gray", ylim = c(0,7), xlab = expression(theta),
     main = "Beta(1,1) Prior vs Beta(3,22) Posterior Distribution",
     ylab = expression(paste(pi, "(", theta, ")", sep="")))
lines(xx,posterior,type = "l", col = "black")
legend(.8,7,legend = c("Beta(1,1)", "Beta(3,22)"),col=c("gray", "black"), lty = 1,cex=0.8)
```

(c) Find a 99.5% (posterior) credible interval for this proportion.

```{r}
creds <- qbeta(c(0.0025, 0.9975),alphastar,betastar)
paste("We are 99.5% sure that the posterior probability of the proportion of Harvard graduates who would answer this question correctly is between",round(creds[1], 4),"and",round(creds[2],4),".")
```

(6) For all parts of this question, you must get the answer in two ways: first, using Monte Carlo to estimate the requested quantities, and then exactly calculating the answer (using built-in R functions, if necessary). Make sure (as always) to include your R code.

(a) P(Y < −13) if Y ∼ N(µ = −10, σ2 = 4).

Monte Carlo

```{r}
library(dplyr)
# If we were actually looking for P(Y < -13) out of a monte-carlo distribution
mean(rnorm(10000,-10,4)  < -13)
```

Exact Answer

```{r}
pnorm(-13,-10,4)
```

(b) 90% (posterior) credible interval for θ if Y |θ ∼ Binomial(42, θ), θ ∼ Beta(5.8, 7.1), and y = 19.

Monte Carlo

```{r}
n <- 42 ; y <- 19 ; alpha <- 5.8 ; beta <- 7.1
alphastar <- alpha + y
betastar <- n - y + beta
theta <- rbeta(10000,alphastar,betastar)
quantile(theta, c(0.05, 0.95))
```

Exact Answer

```{r}
lower <- qbeta(.05,alphastar,betastar)
upper <- qbeta(.95,alphastar,betastar)
c("5%" = lower, "95%" = upper)
```

(c) E(X) if X ∼ Binomial(109, .63)

Monte Carlo means

```{r}
replicate(1000,mean(rbinom(100,109,.63))) %>% mean
```

Exact Answer

```{r}
109*.63
```


