---
title: "Homework 2"
output: html_notebook
---

(1) For this question, we are working with the following probability mass function (pmf).

$$
f(x) =
\frac{7!}{x!(7 − x)!}(.6)^x(.4)^(7-x) 
\cdots
{\substack{x∈\{0,1,2,3,4,5,6,7\}}}
$$

(a) Plot the pmf using R, and make sure to include the graphic and your R code in your answer. The axis
labels and overall chart title should be descriptive–make the plot look nice and be self-explanatory.

```{r, echo = T}
k <- seq(0,7)
plot(k,dbinom(k,size=7,prob=.6),type="h", 
     lwd = 3, xlab = "x",
     main = "Binomial (n = 7, p = .6) Distribution",
     ylab = "F(x|n = 7,p = .6) : Probability Mass", col = "steel blue")



```

(b) What is the mean of X’s distribution?

```{r}
# mean = np
n = 7
p = .6
n*p
sum(dbinom(k,7,.6) * k) # or..
```

(c) What is the support of X’s distribution?

```{r}
k
```

(2) For this question, we are working with the following probability density function (pdf).

$$
f(x) = (1/100) exp(x/100) 
\cdots 
{\substack{(x<0)}}
$$

(a) Plot the pdf using R, and make sure to include the graphic and your R code in your answer. Also, make sure that you do NOT call the y-axis the probability, but rather the probability density (or density, for short).

```{r}
x <- seq(0,-200,-.1)
fx <- (1/100)*exp(x/100)

plot(x,fx, type = "l", lwd = 3, ylim = c(0,.010),
     ylab = "F(x|λ = 1/100) : Probability Density", col = "forest green",
     main ="PDF of X (1/λ = 100) " )
```

(b) What is the support of X’s distribution?

```{r}
"x < 0"
```

(3) Identify the following by using the appropriate R code. In each part below, the assumption is that Y has the Normal distribution with mean 20 and variance=42=16. In the future, I would abbreviate this by saying..

$$
Y ∼ N(20, 4^2)
$$

(a) What is the standard deviation of Y ? (No R code necessary for this part)

```{r}
"4"
```

(b) P(Y < 28.4).

```{r}
pnorm(28.4,20,4)
```


(c) P(Y > 14).

```{r}
1-pnorm(14,20,4)
```

(d) P(17.4 < Y ≤ 23.9).

```{r}
pnorm(23.9,20,4) - pnorm(17.4,20,4) 
```

(e) The first quartile (i.e., 25th percentile) of Y’s distribution.

```{r}
qnorm(.25,20,4)
```


(f) The probability density at the distribution’s mode (i.e., what is f(20), because’s the distribution’s peak is at the value y = 20).

```{r}
dnorm(20,20,4)
```

