---
title: "Homework 8"
output: 
  html_notebook:
    toc: true
    toc_float: true
    theme: lumen
---

### (1)

You have gathered mean arterial pressures from ten subjects currently being treated for high blood pressure with a specific drug regimen. Here are the data:

```{r}
mapr <- c(97.6, 117.7, 126.7, 111.0, 106.7, 108.4, 118.9, 108.6, 119.1, 106.2)
```

You assume these data are (conditionally) normally distributed.

#### (a) Compute the log of the likelihood with µ = 112 and σ^2 = (8.4)^2

```{r}
logl <- function(µ,σ2,y) {
  n <- length(y)
  a <- -(1/2) * n * log(2 * pi)
  b <- -(1/2) * n * log(σ2)
  c <- -(1 / (2 * σ2)) * sum((y - µ)^2)
  a + b + c
}

logl(112,8.4^2,mapr)
log(prod(dnorm(mapr,112,8.4)))
```

#### (b) Compute the log of the likelihood with µ = ybar and σ2 = (1/n)sum(yi-ybar)^2 where ybar = (1/n) sum (yi).

```{r}
logl(mean(mapr),var(mapr),mapr)
log(prod(dnorm(mapr,mean(mapr),sd(mapr))))
```

#### (c) Now plot the log of the likelihood for values of µ ranging from 110 to 114 in increments of .05, and with σ2 fixed at σ2 = (1/n)sum(yi-ybar)^2 where ybar = (1/n) sum (yi).

```{r}
likelihood <- sapply(seq(110,114,.05),function(i) logl(i,var(mapr),mapr))

plot(seq(110,114,.05),likelihood, type = 'l', main = "Likelihood of µ from 110 to 114",
     xlab = expression(theta), ylab = expression(paste(pi,"(", theta,")", sep = "")))
points(seq(110,114,.05)[which.max(likelihood)],max(likelihood), col = "red", pch = 19)
#abline(v = seq(110,114,.05)[which.max(likelihood)], col = "red", h = max(likelihood))
rm(list=ls())
```

<hr>

### (2) 

Here is the total serum cholesterol for eight urban residents of Guatemala:

```{r}
cholest.u <- c(197, 199, 214, 217, 222, 227, 228, 234)
```

The population variance of serum cholesterol measurements for urban residents of Guatemala is assumed to be 260.
The prior belief about the serum cholesterol of residents of Guatemala is that the population mean, µ(urban), should
have a mean of 150 with a variance of 90 (and the prior will be a normal distribution).

#### (a) What is the posterior distribution of the mean, µ(urban), in this case if it is assumed that the data are conditionally iid with a normal distribution?

```{r}
# Prior N(150,90)
m <- 150 ; v <- 90
# posterior params.
starval <- function(m,v,y,sig2) {
  n <- length(y)
  ybar <- mean(y)
  mstar <- (n*v*ybar + sig2*m)/(n*v + sig2)
  s2star <- v*sig2/(n*v + sig2)
  
  c("m*" = mstar, "s2*" = s2star)
}
post.u <- starval(m,v,cholest.u,260)
paste("Posterior is Normal(",round(post.u[1],3),",",round(post.u[2],3),")", sep = "")
```

<i> Normal Posterior distribution </i>

#### (b) Plot the posterior distribution from part (a).

```{r}
xax <- seq(170,230)
plot(xax,dnorm(xax,post.u[1],sqrt(post.u[2])),col='steel blue', type = 'l',
     main = expression("Posterior of µ"[urban]),
     xlab = expression(theta), ylab = expression(paste(pi,"(", theta,")", sep = "")))
```

<hr>

### (3) 

Here is the total serum cholesterol for ten rural residents of Guatemala:

```{r}
cholest.r <- c(139, 142, 143, 144, 145, 148, 155, 162, 171, 181)
```

The population variance is assumed to be 270. The prior belief about the serum cholesterol of residents of Guatemala
is that the population mean, µ(rural), should have a mean of 150 with a variance of 90 (and be normally distributed).

#### (a) What is the posterior distribution of the mean, µ(rural), in this case if it is assumed that the data are conditionally
iid with a normal distribution?

```{r}
post.r <- starval(m,v,cholest.r,270)
paste("Posterior is Normal(",round(post.r[1],3),",",round(post.r[2],3),")", sep = "")
```

#### (b) Plot the posterior distribution from part (a).

```{r}
xax <- seq(130,175)
plot(xax,dnorm(xax,post.r[1],sqrt(post.r[2])),col='steel blue', type = 'l',
     main = expression("Posterior of µ"[rural]),
     xlab = expression(theta), ylab = expression(paste(pi,"(", theta,")", sep = "")))
```

<hr>

### (4) 

Extending the previous two questions, suppose that the urban data are also assumed to be independent of the rural
data. Furthermore, assume that a priori µ(rural) is independent of µ(urban). With these assumptions, µ(rural) and µ(urban)
will also be independent of each other a posteriori.

#### (a) What is the posterior distribution of µ(urban) − µ(rural)?

```{r}
u <- starval(m,v,cholest.u,260); r <- starval(m,v,cholest.r,270)
paste("Normal(",round(u[1]-r[1],4),",",round(u[2]+r[2],4),")", sep = "")
```

#### (b) *Plot the prior and the posterior distribution of µ(urban) − µ(rural) on the same graph.*

```{r}
# Prior µ: E(urban - rural) = 0 ; Prior σ: V(urban - rural) = 180
xax <- seq(-100,100,length.out = 1001)
plot(xax,dnorm(xax,0,sqrt(180)), ylim = c(0,.07), type = "l", col = "steel blue",
     main = expression("Prior & Posterior Distribution of µ"[urban]*" - µ"[rural]),
     xlab = expression(theta),ylab = expression(paste(pi,"(", theta,")", sep = "")))
lines(xax,dnorm(xax,u[1]-r[1],sqrt(u[2]+r[2])), col = "forest green")
```

#### (c) State and interpret in the context of this problem the (central) 95% posterior credible interval for µ(urban) − µ(rural).

```{r}
difs <- qnorm(c(0.025,.975), u[1]-r[1], sqrt(u[2]+r[2]))
paste("After considering the prior and our data, the mean difference in serum cholesterol measurements between urban and rural residents of Guatemal is between ",round(difs[1],4)," and ",round(difs[2],4)," with 95% posterior probability.", sep = "" )
```

<hr>


### (5) 

*GIVE THIS A TRY!!*

The number of immature red tail hawks sitting on fence posts or power poles along a mile-long stretch of rural
highway can reasonably be assumed to follow a Poisson distribution. However, we will be comparing data six years
apart (in 2003 and then in 2009), and we suspect there might be differences in behavior between 2003 and 2009. Thus,
we will assume the observations from 2003 follow the Poisson(θ 2003) distribution, and the observations from 2009
follow the Poisson(θ 2009) distribution. We’ll also assume the 2003 data are independent of the 2009 data. Assume
that a priori θ2003 and θ2009 are independent of each other.

#### (a) Along a 10 mile stretch of highway south of Saratoga Springs in 2003, the count of hawks per mile was 

```{r}
hawks.2003 <- c(0, 2, 0, 0, 0, 4, 3, 0, 0, 1) 
```

What is the expected value of the posterior distribution for θ2003 if the prior is a gamma distribution with shape=0.2 and rate=0.2?

```{r}
astar_2003 <- 0.2 + sum(hawks.2003)
bstar_2003 <- 0.2 + length(hawks.2003)
"The expected value of the posterior distribution of θ_2003 is %s" %>%
  sprintf(astar_2003/bstar_2003)

sprintf("The expected value of the posterior distribution of θ_2003 is %s",astar_2003/bstar_2003)
```

#### (b) Along that same stretch of highway in 2009, the counts were

```{r}
hawks.2009 <- c(1,1,3,3,2,2,2,4,0,3)
```

What is the expected value of the posterior distribution for θ2009 if the same prior is used as in the previous
problem (i.e., Gamma(0.2,0.2)?

```{r}
astar_2009 <- 0.2 + sum(hawks.2009)
bstar_2009 <- 0.2 + length(hawks.2009)
"The expected value of the posterior distribution of θ_2009 is %s" %>%
  sprintf(round(astar_2009/bstar_2009,4))
```

#### (c) What is the expected value of the posterior distribution of d = θ2009 − θ2003?

```{r}
# MC

"The expected value of the posterior distribution of d is %s" %>%
  sprintf(round(mean(rgamma(10000,astar_2009,bstar_2009) - rgamma(10000,astar_2003,bstar_2003)),4))

```


#### (d) What is the 95% posterior credible interval for d?

```{r}
hawks <- quantile(rgamma(100000,astar_2009,bstar_2009) - rgamma(100000,astar_2003,bstar_2003),c(.025,.975))
paste("After considering the prior and our data, the mean difference in immature red tail hawks along a mile long stretch of rural highway is between ", round(hawks[1],4), " and ", round(hawks[2],4)," with 95% posterior probability" , sep = "" )
```


#### (e) Plot the Monte-Carlo estimated posterior distribution of d = θ2009−θ2003 by using the density and plot functions
in R.

```{r}
obj <- rgamma(100000,astar_2009,bstar_2009) - rgamma(100000,astar_2003,bstar_2003)
plot(density(obj), main = expression("MC Estimated Density of θ"[2009]*"−θ"[2003]),
     xlab = expression(theta), ylab = expression(paste(pi,"(", theta,")", sep = "")))
rm(list = ls())
```

