---
title: "Stat_451_Notes"
output: word_document
---

Notes:
--xbar has a sampling distribution
--confidence interval--called that because we are confident in our method and that it would work if repeated many times
--Parameters are fixed but unknown for frequentists
--post(parameters|data) oc likelihood(data|parameters) * prior(parameters)
--As frequentists, we'll assume the data is normal (in likelihood function)
--As Bayesians, we assume parameters in the prior are random variables and have a density while parameters are fixed in frequentists
--Pr(A^B) = Pr(A) * Pr(B) iff A and B are independent
--Pr(A^B) = Pr(A|B) * Pr(B) = Pr(B|A) * Pr(A)
--Pr(A|B) = Pr(B|A) * Pr(A) / Pr(B/.)

```{r}
postmu <- function(data, sigma2, prior_mu, prior_sig2) {
  xbar <- mean(data)
  n <- length(data)
  tau2 <- 1/sigma2
  prior_tau2 <- 1/prior_sig2
  post_mu <- (n*tau2*xbar + prior_tau2*prior_mu) / (n*tau2 + prior_tau2)
  post_prec <- n*tau2 + prior_tau2
  post_var <- 1/post_prec
  post_sd <- sqrt(post_var)
  out <- c(post_mu, post_sd, post_var)
  return(out)
}
```

```{r}
update_sig2 <- function(data, a, s, mu) {
  n <- length(data)
  SSx <- sum((data - mu)^2)
  post_shape <- n/2 + a
  post_scale <- 2*s/(s*SSX+2)
  c(post_shape, post_scale)
}
```

```{r}
igpdf <- function(x, shape, rate) {
  ((rate^shape) / gamma(shape)) * (x^(-shape - 1)) * exp(-rate/x)
}
```
Good starting values for prior mean and variance are mean(data) and var(data)


Gibbs Sampler
```{r}
gibbs_samp <- function(data, prior_mu, prior_var, shape, rate) {
  out <- matrix(0, 1000, 2)
  out[1,1] <- mean(data)
  out[1,2] <- var(data)
  
  for (i in 2:1000) {
    # get posterior values for parameters for normal mean update -- post ~ N(m,v)
    out[i,1] <- rnorm(1, out[i-1, 1], sqrt(out[i-1, 2]))
    # get posterior values for IG update (shape, rate)
    out[i,2] <- 1/rgamma(1, shape, rate)
  }
}
```

```{r}
set.seed(5)
dat1 <- rbinom(20, 1, .2)
dat1

.3 + c(-1,1) * 1.96 * sqrt(.3*.7/20)
```

```{r}
curve(dbeta(x, 7, 15), from = 0, to = 1)

qbeta(c(0.025, 0.975), 7, 15)

curve(dbeta(x, 6, 14), from = 0, to = 1)
curve(dbeta(x, 12, 28), from = 0, to = 1)

qbeta(c(0.025, 0.975), 12, 28)
```

The likelihood now dominates the prior if the number of data points is much greater than the prior values.

Credible interval = posterior probability interval


```{r}
g <- function(data, theta, a, b) {
  theta^(sum(data)+a-1)*(1-theta)^(length(data)-sum(data)+b-1)
}

lng <- function(data, theta, a, b) {
  (sum(data)+a-1)*log(theta) + (length(data)-sum(data)+b-1)*log(1-theta)
}

metBern <- function(data, pra, prb, loops, candsig) {
  cntr <- 0
  out <- NULL
  out[1] <- .5
  for (i in 2:loops) {
    out[i] <- out[i-1]
    cand <- rnorm(1, out[i-1], candsig)
    if (cand > 0 & cand < 1) {
      r <- lng(data, cand, pra, prb) - lng(data, out[i-1], pra, prb)
      if (r > log(runif(1,0,1))) {
        out[i] <- cand
        cntr <- cntr + 1
      }
    }
  }
  return(list(out = out, cntr = cntr))
}
```

Counter (cntr) should be somewhere between .15 and .4 for a metropolis sampler.


Jags
```{r}
library(rjags)
library(R2jags)

mdl <- "
  model {

   for (i in 1:N) {
     y[i] ~ dbern(theta)
   }
   theta ~ dbeta(1,1)
  }

"

writeLines(mdl, 'bern.txt')
set.seed(5)
y <- rbinom(10, 1, .3)
N <- length(y)
data.jags <- c('y', 'N')
parms <- c('theta')
bern.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms,
                 model.file = 'bern.txt', n.chains = 5, n.iter = 5000, n.burnin = 1000,
                 n.thin = 1)
```

Rhat comes from having more than 1 chain--An Rhat of 1 says that all chains are converging to the same place.
Deviance: -2 * log likelihood
Information criteria: -2 * log likelihood + penalty
Deviance Information Criteria: 
p_d = variance(deviance) / 2 (We want it to be about the number of parameters there are)
DIC = mean(deviance) + p_d (penalty)

Another way to compute p_d = mean(deviance) - deviance(mean(theta))

autocorr.diag() allows us to look at autocorrelation by lagging the values to comparing. Takes an as.mcmc() object.--low, reduce by thinning

effectiveSize() shows if samples are independent or not by if the number of draws is close to the output of the function.--above 5000

raftery.diag() should be used on mcmc object of all chains given. Tells accuracy in tail. Look at dependence factor and if it's about 1 for theta then it's pretty accurate. If the number gets above 5, you shouldn't be using chains for posterior probability intervals--use mean. 2 or 3 is not good either. Change this number by thinning.

gelman.diag() looks at beginning and end of chains--similar to Rhat

Also plot trace plots to see that it looks reasonable

```{r}
library('R2jags')
# Remember you put in precisions and not variances in jags
mdl <- "
  model {
    for (i in 1:33) {
      y[i] ~ dnorm(mu[tmt[i]], 1/vr[tmt[i]])
    }
  # vr ~ dunif(0, 4000)
    for (i in 1:2) {
      mu[i] ~ dnorm(0, 0.000001)
      vr[i] ~ dunif(0, 5000)
  # prec[i] <- 1/vr[i]
    }
  }
"
writeLines(mdl, "twogroups.txt")
dat <- read.table('01twogroups.dat')
tmt <- dat$V1
y <- dat$V2

data.jags <- c('tmt', 'y')
parms <- c('mu', 'vr')

twogroups.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms,
                      model.file = 'twogroups.txt', n.iter = 50000, n.burnin = 10000,
                      n.chains = 5, n.thin = 5)
```

```{r}
# Some analysis after running simulation
sims <- as.mcmc(twogroups.sim)
chains <- as.matrix(sims)
sims2 <- as.mcmc(chains) # Combine all chains into 1 chain

# Standard diagnostics
raftery.diag(sims2)
effectiveSize(sims2)
autocorr.diag(sims2) # You can reduce auto-correlation by thinning

# Look at ratio of variances
ratvar <- chains[,4] / chains[,5]
plot(density(ratvar))
mean(ratvar > 1)
plot(sims2)
plot(chains[,4], type = 'l')
# If you have more knowledge of the variance, you probably want to use a gamma prior rather than a uniform prior
curve(dgamma(x, 20, 1/50), from = 0, to = 4000)
```
Try model defining different variances
```{r}
library('R2jags')
# Remember you put in precisions and not variances in jags
mdl <- "
  model {
    for (i in 1:33) {
      y[i] ~ dnorm(mu[tmt[i]], 1/vr[tmt[i]])
    }
   vr[1] ~ dgamma(20, .02)
   vr[2] ~ dgamma(10, .1)
    for (i in 1:2) {
      mu[i] ~ dnorm(0, 0.000001)
      #vr[i] ~ dunif(0, 5000)
  # prec[i] <- 1/vr[i]
    }
  }
"
writeLines(mdl, "twogroups.txt")
dat <- read.table('01twogroups.dat')
tmt <- dat$V1
y <- dat$V2

data.jags <- c('tmt', 'y')
parms <- c('mu', 'vr')

twogroups.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms,
                      model.file = 'twogroups.txt', n.iter = 50000, n.burnin = 10000,
                      n.chains = 5, n.thin = 5)

plot(density(chains[,4]), xlim = c(0, 2000), ylim = c(0, 0.01))
lines(density(chains[,5]), col = 'red')

mudif <- chains[,3]-chains[,2]
plot(density(mudif))
mean(mudif > 0) # Posterior probability that the first group has a smaller mean than the second group
quantile(mudif, c(0.025, 0.975))
```


Jags One-Way Anova

```{r}
library(R2jags)
mdl <- "
  model {
    for (i in 1:28) {
      response[i] ~ dnorm(mu[tmt[i]], prec)
    }
    prec <- 1/vv
    vv ~ dgamma(2, .1)
    for (i in 1:4) {
      mu[i] ~ dnorm(15, .0001)
    }
  }
"

writeLines(mdl, 'anova.txt')
dat <- read.table('04anova.dat')
tmt <- dat$V1
response <- dat$V2

data.jags <- c('tmt', 'response')
parms <- c('mu', 'vv')

anova.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms,
                      model.file = 'anova.txt', n.iter = 42000, n.burnin = 2000,
                      n.chains = 5, n.thin = 8)
```
```{r}
mu1vmu2 <- chains[,2] - chains[,3]
mu1vmu3 <- chains[,2] - chains[,4]
mu1vmu4 <- chains[,2] - chains[,5]
mu2vmu3 <- chains[,3] - chains[,4]
mu2vmu4 <- chains[,3] - chains[,5]
mu3vmu4 <- chains[,4] - chains[,5]

plot(density(mu1vmu2), type = 'l')
plot(density(mu1vmu3), type = 'l')
```
Test Interaction
```{r}
intact <- chains[,2] - chains[,3] - chains[,4] + chains[,5]
plot(density(intact))
mean(intact < 0)
```
Marginal Effects
```{r}
mainirg <- .5*chains[,2] + .5*chains[,3] - .5*chains[,4] - .5*chains[,5]
plot(density(mainirg))
mainfert <- .5*chains[,2] + .5*chains[,4] - .5*chains[,3] - .5*chains[,5]
plot(density(mainfert))
```

